---
title: "Team 2 File"
author: "Michael Egle, John Chandara, Jaydon Cobb, Anthony Song"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(readxl)
library(ggplot2)
```

#Michael Egle Analysis
```{r}
case <- read_csv(url('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv'))

death <- read_csv(url('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv'))

rec <- read_csv(url('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv'))

mob <- read_csv('applemobilitytrends-2020-04-24.csv')

nyc <- read_csv('nyc_covid.csv') #from nyc.gov

counties <- read_csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
```

```{r}
mob <- mob %>%
  pivot_longer(`2020-01-13`:`2020-04-24`,
               names_to = "date",
               values_to = "percent_change")
```

```{r}
counties <- counties %>%
  mutate(date = ymd(date))

mob <- mob %>%
  mutate(date = ymd(date))
```

Break up the county data set into four datasets for our cities of interest
```{r}
nyc <- counties %>%
  filter(county == "New York City") %>%
  select(-fips)

mia <- counties %>%
  filter(county == "Miami-Dade") %>%
  select(-fips)

chi <- counties %>%
  filter(county == "Cook", state == "Illinois") %>%
  select(-fips)

sf <- counties %>%
  filter(county == "San Francisco") %>%
  select(-fips)
```


Use inner_join and some date manipulations to get the cases from the previous day
```{r}
nyc2 <- nyc
mia2 <- mia
chi2 <- chi
sf2 <- sf

nyc <- nyc %>%
  mutate(day_before = date - 1)

mia <- mia %>%
  mutate(day_before = date - 1)

chi <- chi %>%
  mutate(day_before = date - 1)

sf <- sf %>%
  mutate(day_before = date - 1)

nyc2 <- nyc2 %>%
  select(c('date', 'deaths'))

mia2 <- mia2 %>%
  select(c('date', 'deaths'))

chi2 <- chi2 %>%
  select(c('date', 'deaths'))

sf2 <- sf2 %>%
  select(c('date', 'deaths'))

colnames(nyc2) <- c("day_before", "deaths_before")
colnames(mia2) <- c("day_before", "deaths_before")
colnames(chi2) <- c("day_before", "deaths_before")
colnames(sf2) <- c("day_before", "deaths_before")

nyc <- nyc %>%
  inner_join(nyc2)

mia <- mia %>%
  inner_join(mia2)

chi <- chi %>%
  inner_join(chi2)

sf <- sf %>%
  inner_join(sf2)
```

```{r}
nyc <- nyc %>%
  mutate(new_deaths = deaths - deaths_before)

mia <- mia %>%
  mutate(new_deaths = deaths - deaths_before)

chi <- chi %>%
  mutate(new_deaths = deaths - deaths_before)

sf <- sf %>%
  mutate(new_deaths = deaths - deaths_before)
```


```{r}
nyc_travel <- mob %>%
  filter(region == "New York City")

mia_travel <- mob %>%
  filter(region == "Miami")

chi_travel <- mob %>%
  filter(region == "Chicago")

sf_travel <- mob %>%
  filter(region == "San Francisco - Bay Area")
```


Put together graph for NYC
```{r}
nyc_travel <- nyc_travel %>%
  filter(date > ymd("2020/2/29"))
```

```{r}
nyc_travel_viz <- nyc_travel %>%
  ggplot(aes(x = date, y = percent_change, group = transportation_type,
             color = transportation_type)) +
  geom_line() +
  geom_smooth() +
  theme(legend.position = "none") +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 8, label = "Transit",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 35, label = "Walking",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 55, label = "Driving",
           size = 3) +
  geom_vline(xintercept = ymd("2020/3/20"), linetype = "dashed") +
  annotate(geom = "label", x = ymd("2020/3/20"), y = 125,
           label = "Stay At Home Order Put Into Effect 3/20", size = 3) +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  scale_y_continuous(limits = c(0, 200)) +
  labs(title = "New York City",
       subtitle = "26,403 people per square mile",
       y = "Percent of Average Movement") +
  theme(axis.title.x = element_blank())
```

```{r}
nyc_deaths_viz <- nyc %>%
  ggplot(aes(x = date, y = new_deaths)) +
  geom_histogram(stat = "identity") +
  geom_vline(xintercept = ymd("2020/3/20"), linetype = "dashed") +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  geom_smooth() +
  labs(caption = "Movement data from Apple Maps | Deaths data from New York Times",
       x = "Date",
       y = "New Deaths") +
  annotate(geom = "text", x = ymd("2020/4/12"), y = 1150,
           label = "11,157 total deaths as of 4/24", size = 3)
```

```{r}
ggarrange(nyc_travel_viz, nyc_deaths_viz,
          nrow = 2)
```

Now let's put together Miami's graphic
```{r}
mia_travel <- mia_travel %>%
  filter(date > ymd("2020/2/29"))
```

```{r}
mia_travel_viz <- mia_travel %>%
  ggplot(aes(x = date, y = percent_change, group = transportation_type,
             color = transportation_type)) +
  geom_line() +
  geom_smooth() +
  theme(legend.position = "none") +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 25, label = "Transit",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 40, label = "Walking",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 55, label = "Driving",
           size = 3) +
  geom_vline(xintercept = ymd("2020/3/26"), linetype = "dashed") +
  annotate(geom = "label", x = ymd("2020/3/26"), y = 140,
           label = "Stay At Home Order Put Into Effect for Miami-Dade County 3/26",
           size = 3) +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  scale_y_continuous(limits = c(0, 200)) +
  labs(title = "Miami",
       subtitle = "12,139 people per square mile",
       y = "Percent of Average Movement") +
  theme(axis.title.x = element_blank())

```

```{r}
mia_deaths_viz <- mia %>%
  ggplot(aes(x = date, y = new_deaths)) +
  geom_histogram(stat = "identity") +
  geom_vline(xintercept = ymd("2020/3/26"), linetype = "dashed") +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  geom_smooth() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(caption = "Movement data from Apple Maps | Deaths data from New York Times",
       x = "Date",
       y = "New Deaths") +
  annotate(geom = "text", x = ymd("2020/4/12"), y = 70,
           label = "287 total deaths as of 4/24", size = 3)

```

```{r}
ggarrange(mia_travel_viz, mia_deaths_viz,
          nrow = 2)
```

Now let's do Chicago
```{r}
chi_travel <- chi_travel %>%
  filter(date > ymd("2020/2/29"))
```

```{r}
chi_travel_viz <- chi_travel %>%
  ggplot(aes(x = date, y = percent_change, group = transportation_type,
             color = transportation_type)) +
  geom_line() +
  geom_smooth() +
  theme(legend.position = "none") +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 15, label = "Transit",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 42, label = "Walking",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 69, label = "Driving",
           size = 3) +
  geom_vline(xintercept = ymd("2020/3/20"), linetype = "dashed") +
  annotate(geom = "label", x = ymd("2020/3/20"), y = 170,
           label = "Stay At Home Order Put Into Effect 3/20", size = 3) +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  scale_y_continuous(limits = c(0, 200)) +
  labs(title = "Chicago",
       subtitle = "11,943 people per square mile",
       y = "Percent of Average Movement") +
  theme(axis.title.x = element_blank())

```

```{r}
chi_deaths_viz <- chi %>%
  ggplot(aes(x = date, y = new_deaths)) +
  geom_histogram(stat = "identity") +
  geom_vline(xintercept = ymd("2020/3/20"), linetype = "dashed") +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  geom_smooth() +
  labs(caption = "Movement data from Apple Maps | Deaths data from New York Times",
       x = "Date",
       y = "New Deaths") +
  scale_y_continuous(limits = c(0, 125)) +
  annotate(geom = "text", x = ymd("2020/4/12"), y = 112,
           label = "1,220 total deaths as of 4/24", size = 3)
```

```{r}
ggarrange(chi_travel_viz, chi_deaths_viz,
          nrow = 2)
```

And finally San Francisco
```{r}
sf <- sf %>%
  filter(date > ymd("2020/2/29"))
```

```{r}
sf_travel_viz <- sf_travel %>%
  ggplot(aes(x = date, y = percent_change, group = transportation_type,
             color = transportation_type)) +
  geom_line() +
  geom_smooth() +
  theme(legend.position = "none") +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 10,
           label = "Transit",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 32,
           label = "Walking",
           size = 3) +
  annotate(geom = "text", x = ymd("2020/4/17"), y = 52,
           label = "Driving",
           size = 3) +
  geom_vline(xintercept = ymd("2020/3/16"), linetype = "dashed") +
  annotate(geom = "label", x = ymd("2020/3/16"), y = 125,
           label = "Stay At Home Order Put Into Effect 3/16",
           size = 3) +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24")))  +
  scale_y_continuous(limits = c(0, 200)) +
  labs(title = "San Francisco",
       subtitle = "17,246 people per square mile",
       y = "Percent of Average Movement") +
  theme(axis.title.x = element_blank())
```

```{r}
sf_deaths_viz <- sf %>%
  ggplot(aes(x = date, y = new_deaths)) +
  geom_histogram(stat = "identity") +
  geom_vline(xintercept = ymd("2020/3/16"), linetype = "dashed") +
  scale_x_date(limits = c(ymd("2020-03-01", "2020-04-24"))) +
  geom_smooth() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(caption = "Movement data from Apple Maps | Deaths data from New York Times",
       x = "Date",
       y = "New Deaths") +
  annotate(geom = "text", x = ymd("2020/4/12"), y = 50,
           label = "22 total deaths as of 4/24", size = 3)
```

```{r}
ggarrange(sf_travel_viz, sf_deaths_viz,
          nrow = 2)
```



#Anthony Song Analysis 
Read in the data
```{r}
mob <- read_csv('applemobilitytrends-2020-04-24.csv')
```
pivot longer and month year date
```{r}
mob <- mob %>%
  pivot_longer(`2020-01-13`:`2020-04-24`,
               names_to = "date",
               values_to = "percent_change")
mob <- mob %>%
  mutate(date = ymd(date))
```
Filter some low rate countries
```{r}
mob1 <- mob %>%
  filter(transportation_type == "walking")
# Very low death rate per million people
indo <- mob1 %>%
  filter(region == "Indonesia")
cam <- mob1 %>%
  filter(region == "Cambodia")
kor <- mob1 %>%
  filter(region == "Republic of Korea")
tai <- mob1 %>%
  filter(region == "Taiwan")
tha <- mob1 %>%
  filter(region == "Thailand")
low_rate <- rbind(cam, tha)
```
First Visual
```{r}
reg_viz3 <- low_rate%>%
  ggplot(aes(x = date, y = percent_change))+geom_line()+facet_wrap(~region)+geom_smooth()+ylim(0,200)
reg_viz3+labs(x = "Date", y= "Percent Mobility Relative to Baseline", title = "Mobility of 2 Low Death Rate Regions")+geom_hline(yintercept = 125, linetype = "dashed", color = "red")+
  annotate(geom = "label", x = ymd("2020/4/10"), y = 125,
           label = "125% Mobility",
           size = 3)

```
#These Countries have very low death rate per million, their rate small rate of change and activity agree with this. 
```{r}
# Rate of change of Indonesia
qplot(date,percent_change,data=indo) + stat_smooth(aes(outfit=fit_indo<<-..y..))+annotate("rect", ymin = 25, ymax = 100, xmin = ymd("2020/3/1"), xmax = ymd("2020/4/1"), alpha = .2)+labs(title = "Indonesia Rate of Change")
fit_indo <- data.frame(fit_indo)
fit_indo$day<- seq.int(nrow(fit_indo))
vc_indo <- c(max(fit_indo$fit_indo), min(fit_indo$fit_indo) )
max_min_indo <- fit_indo[fit_indo$fit_indo %in% vc_indo,]
print("Indonesia ROC")
(max_min_indo[2,1] - max_min_indo[1,1]) / (max_min_indo[2,2] -max_min_indo[1,2])

# rate of change of Cambodia
qplot(date,percent_change,data=cam) + stat_smooth(aes(outfit=fit_cam<<-..y..))
fit_cam <- data.frame(fit_cam)
fit_cam$day<- seq.int(nrow(fit_cam))
vc_cam <- c(max(fit_cam$fit_cam), min(fit_cam$fit_cam) )
max_min_cam <- fit_cam[fit_cam$fit_cam %in% vc_cam,]
print("Camboia ROC")
(max_min_cam[2,1] - max_min_cam[1,1]) / (max_min_cam[2,2] -max_min_cam[1,2])

# rate of change of Soutn korea
qplot(date,percent_change,data=kor) + stat_smooth(aes(outfit=fit_kor<<-..y..))
fit_kor <- data.frame(fit_kor)
fit_kor$day<- seq.int(nrow(fit_kor))
vc_kor <- c(max(fit_kor$fit_kor), min(fit_kor$fit_kor) )
max_min_kor <- fit_kor[fit_kor$fit_kor %in% vc_kor,]
print("South Korea ROC")
(max_min_kor[2,1] - max_min_kor[1,1]) / (max_min_kor[2,2] -max_min_kor[1,2])

# rate of change of Taiwan
qplot(date,percent_change,data=tai) + stat_smooth(aes(outfit=fit_tai<<-..y..))
fit_tai <- data.frame(fit_tai)
fit_tai$day<- seq.int(nrow(fit_tai))
vc_tai <- c(max(fit_tai$fit_tai), min(fit_tai$fit_tai) )
max_min_tai <- fit_tai[fit_tai$fit_tai %in% vc_tai,]
print("Taiwan ROC")
(max_min_tai[2,1] - max_min_tai[1,1]) / (max_min_tai[2,2] -max_min_tai[1,2])

# rate of change of Thailand
qplot(date,percent_change,data=tha) + stat_smooth(aes(outfit=fit_tha<<-..y..))
fit_tha <- data.frame(fit_tha)
fit_tha$day<- seq.int(nrow(fit_tha))
vc_tha <- c(max(fit_tha$fit_tha), min(fit_tha$fit_tha) )
max_min_tha <- fit_tha[fit_tha$fit_tha %in% vc_tha,]
print("Thailand ROC")
(max_min_tha[2,1] - max_min_tha[1,1]) / (max_min_tha[2,2] -max_min_tha[1,2])
```


Regions
```{r}
#High death rate per million people. 
ity <- mob %>%
  filter(region == "Italy")
por <- mob %>%
  filter(region == "UK")
esp <- mob %>%
  filter(region == "Spain")
bel <- mob %>%
  filter(region == "Belgium")
ire <- mob %>%
  filter(region == "Ireland")
usa <- mob %>%
  filter(region == "United States")

non_rec_regions <- rbind( por, esp, ire, bel)
non_rec_regions <- non_rec_regions %>%
  filter(transportation_type == "walking")

reg_viz2 <- non_rec_regions %>%
  ggplot(aes(x = date, y = percent_change))+geom_line()+facet_wrap(~region)+geom_smooth()+ylim(0,200)
reg_viz2+labs(x = "Date", y = "Percent Change in Mobility", title = "Mobility of 4 High Death Rate Regions")+labs(x = "Date", y= "Percent Change in Mobility")+geom_hline(yintercept = 125, linetype = "dashed", color = "red")+annotate(geom = "label", x = ymd("2020/4/10"), y = 125,
           label = "125% Mobility",
           size = 3)

# How does the USA compare???
reg_viz_USA <- usa %>%
  ggplot(aes(x = date, y = percent_change))+geom_line()+facet_wrap(~region)+geom_smooth()+ylim(0,175)
reg_viz_USA+labs(x = "Date", y = "Percent Change in Mobility", title = "Walking Mobility of the USA")+labs(x = "Date", y= "Percent Mobility Relative to Baseline")+geom_hline(yintercept = 125, linetype = "dashed", color = "red")+annotate(geom = "label", x = ymd("2020/4/10"), y = 125,
           label = "125% Mobility",
           size = 3)
```
```{r}
high_low <-rbind(cam, tha, bel, ire)
reg_viz5 <- high_low%>%
  ggplot(aes(x = date, y = percent_change))+geom_line()+facet_wrap(~region)+geom_smooth()+ylim(0,200)+labs(x = "Date", y= "Percent Mobility Relative to Baseline", title = "Contrasting Walking Mobility Data of 4 Regions")+geom_hline(yintercept = 125, linetype = "dashed", color = "red")+ annotate(geom = "label", x = ymd("2020/4/10"), y = 125,label = "125% Mobility",size = 3)
reg_viz5
```

```{r}
# rate of change of italy
qplot(date,percent_change,data=ity) + stat_smooth(aes(outfit=fit_ity<<-..y..))
fit_ity <- data.frame(fit_ity)
fit_ity$day<- seq.int(nrow(fit_ity))
vc_ity <- c(max(fit_ity$fit_ity), min(fit_ity$fit_ity) )
max_min_ity <- fit_ity[fit_ity$fit_ity %in% vc_ity,]
print("Italy ROC")
rate <- c((max_min_ity[2,1] - max_min_ity[1,1]) / (max_min_ity[2,2] -max_min_ity[1,2]))

# rate of change portugal
qplot(date,percent_change,data=por) + stat_smooth(aes(outfit=fit_por<<-..y..))
fit_por <- data.frame(fit_por)
fit_por$day<- seq.int(nrow(fit_por))
vc_por <- c(max(fit_por$fit_por), min(fit_por$fit_por) )
max_min_por <- fit_por[fit_por$fit_por %in% vc_por,]
print("Portugal ROC")
rate <- append(rate, (max_min_por[2,1] - max_min_por[1,1]) / (max_min_por[2,2] -max_min_por[1,2]))

# rate of change spain
qplot(date,percent_change,data=esp) + stat_smooth(aes(outfit=fit_esp<<-..y..))
fit_esp <- data.frame(fit_esp)
fit_esp$day<- seq.int(nrow(fit_esp))
vc_esp <- c(max(fit_esp$fit_esp), min(fit_esp$fit_esp) )
max_min_esp <- fit_esp[fit_esp$fit_esp %in% vc_esp,]
print("Spain ROC")
rate <- append(rate,(max_min_esp[2,1] - max_min_esp[1,1]) / (max_min_esp[2,2] -max_min_esp[1,2]))

# rate of change ireland
qplot(date,percent_change,data=ire) + stat_smooth(aes(outfit=fit_ire<<-..y..))
fit_ire <- data.frame(fit_ire)
fit_ire$day<- seq.int(nrow(fit_ire))
vc_ire <- c(max(fit_ire$fit_ire), min(fit_ire$fit_ire) )
max_min_ire <- fit_ire[fit_ire$fit_ire %in% vc_ire,]
print("Ireland ROC")
rate <- append(rate,(max_min_ire[2,1] - max_min_ire[1,1]) / (max_min_ire[2,2] -max_min_ire[1,2]))

# rate of change Belgium
qplot(date,percent_change,data=bel) + stat_smooth(aes(outfit=fit_bel<<-..y..))
fit_bel <- data.frame(fit_bel)
fit_bel$day<- seq.int(nrow(fit_bel))
vc_bel <- c(max(fit_bel$fit_bel), min(fit_bel$fit_bel) )
max_min_bel <- fit_bel[fit_bel$fit_bel %in% vc_bel,]
print("Belgium ROC")
rate <- append(rate,(max_min_bel[2,1] - max_min_bel[1,1]) / (max_min_bel[2,2] -max_min_bel[1,2]))
rate
```

```{r}
# rate of change of USA
qplot(date,percent_change,data=usa) + stat_smooth(aes(outfit=fit_usa<<-..y..))
fit_usa <- data.frame(fit_usa)
fit_usa$day<- seq.int(nrow(fit_usa))
vc_usa <- c(max(fit_usa$fit_usa), min(fit_usa$fit_usa) )
max_min_usa <- fit_usa[fit_usa$fit_usa %in% vc_usa,]
max_min_usa
(max_min_usa[2,1] - max_min_usa[1,1]) / (max_min_usa[2,2] -max_min_usa[1,2])

```

Read excel
```{r}
dpp <- read_excel("deaths_per_million.xlsx")
```
# clean data
```{r}
colnames(dpp) <- dpp[1, ]
dpp <- dpp[-1, ]
dpp[] <- lapply(dpp, function(x) type.convert(as.character(x)))
dpp <- dpp[complete.cases(dpp),]
# Take absolute value for plotting purposes. 
#dpp$`Rate of decrease in Mobility` <- abs(dpp$`Rate of decrease in Mobility`)
head(dpp)
```

#Run linear regression
```{r}
# Model a polynomial function. 
regressor = lm(dpp$`Deaths per a million people` ~ poly(dpp$`Rate of decrease in Mobility`,2, raw = TRUE), data=dpp)
summary(regressor)
pred_dpp <- data.frame(death = predict(regressor, dpp), rate=dpp$`Rate of decrease in Mobility`)

#Graph with R^2 and p-value 
ggplot(data = dpp, aes(y = `Deaths per a million people` , x = `Rate of decrease in Mobility`)) + 
  geom_point(color='blue') +
  geom_line(color='red',data = pred_dpp, aes(x=rate, y=death))+ labs(x = "Rate of Mobility from Peak to Base (Percent/Day)", title = "Effect of Walking Mobility Rate on COVID-19 Death Rate")+ annotate(geom = "label", x = -1, y = 500,label = "R^2: 0.775, p-value: 0.003",size = 5)+xlim(-4,0)+annotate(geom = "label", x = -1.6, y = 170,label = "USA",size = 4)+annotate(geom = "label", x = -0.8, y = 50,label = "Taiwan",size = 4)+annotate(geom = "label", x = -3.5, y = 450,label = "Spain",size = 4)
```
AS2*